Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex
Pass over related work

At a high level, techniques for automatically distributing fall under a spectrum that ranges from fully automated (but potentially sub-optimal) distribution to fully manual (but, ideally, optimal) distribution (Fig.~\ref{fig:spectrum}).

@@

Several packages expose commands for specifying parallelism on modern \unix{}es---\eg \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}. %  or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.
Different from \sys, their effectiveness is predicated upon explicit invocation and is limited to embarrassingly parallel (and short) programs.
Often, these commands provide options to support an array of special sub-cases---a stark contradiction to the celebrated \unix philosophy.
For example, \tti{parallel}, an official GNU project, contains flags such as \ttt{--skip-first-line}, \ttt{-trim}, and \ttt{--xargs}, that a \unix user would achieve with \ttt{head}, \ttt{sed}, and \ttt{xargs}; it also includes other programs with complex semantics, such as the ability transfer files between computers, separate text files, and parse CSV.
\sys embraces the \unix philosophy and simply attempts to rewrite pipelines more efficiently.

Several shells~\cite{duff1990rc, mcdonald1988support, dagsh:17} add primitives for non-linear pipe topologies---some of which target distribution.
Here too, however, developers are expected to manually rewrite scripts to exploit these primitives.

@@

\heading{Cluster Computing}
Modern analytics frameworks~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} and domain-specific languages~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp} go a step further in terms of automation.
By providing a few carefully-designed primitives, they 
Due to the specialization of these primitives, however, these systems suffer from a lack of generality ;
% and they still require developers to compose their work using the pipelines

They also support a single class of distributability, arguing that is enough 
by making strong assumptions about the nature of the computation---for example, strongly-eventual commutative components that can proceed in parallel.
Developing under these frameworks differs quite significantly from the development of normal (non-distributed) programs.

% Nice intro: http://homepages.inf.ed.ac.uk/bfranke/Publications/pldi121-tournavitis.pdf
\heading{Automated Parallelization and Distribution}
There is a long history of automated parallelization starting from explicit \ttt{DOALL} and \ttt{DOACROSS} annotations~\cite{par1, par2} and continuing with compilers that attempt to automatically extract parallelism~\cite{padua1993polaris,hall1996maximizing}.
These systems operate at a lower level than \sys (\eg that of instructions or loops instead of pipe boundaries) and typically do not exploit runtime information.

% https://ecommons.cornell.edu/bitstream/handle/1813/6508/85-668.pdf?sequence=1
% The growth of the web led to specialized frameworks for massively distributed computation~\cite{mapreduce:08, spark:10, naiad:13}.
% While they Moreover, these systems take advantage of functional purity;
%   for programs that are not data-intensive processing pipelines (\eg web servers), purely functional code is generally responsible for only a small fraction of the program runtime.
\heading{Annotation-based Distribution}
Recent systems for light-touch distribution, Ignis~\cite{ignis:19} and Mozart~\cite{mozart:19}, give developers the ability to express through annotations;

Both work on high-level, managed languages---server-side JavaScript for Ignis and Python for Mozart---that allow runtime transformations without affecting the broader environment under which a program executes.
This is different

Annotations are used by ---complicated by the use of third-party libraries similar to \sys's extensions.
\sys is designed to be used, so it is somewhat more heavy-weight than Ignis' declarative recipes or Mozart's split annotations, but geared toward the deveopers of these components---not the users composing them.
As such..
Their annotations language is not as principled as \sys's---and one that is critical in ensuring that the dataflow analyses preserves the semantics expected by the developer
The side-effects present in the shell dwarf those of functional languages
The shell is also more general.

>>>>

At a high level, techniques for re-writing programs to exploit distribution fall under a wide spectrum that ranges from
  fully manual (but, ideally, optimal) to 
  fully automated (but potentially sub-optimal) distribution   (Fig.~\ref{fig:spectrum}).

@@

Several packages expose commands for specifying parallelism on modern \unix{}es---\eg \ttt{qsub}~\cite{gentzsch2001sun}, \textsc{SLURM}~\cite{yoo2003slurm}, calls to \textsc{GNU} \ttt{parallel}~\cite{Tange2011a}. %  or any manual rewriting~\cite{mapreduce:08, ciel:11, spark:12}.
Different from \sys, their effectiveness is predicated upon explicit invocation and is limited to embarrassingly parallel (and short) programs.
Often, these commands provide options to support an array of special sub-cases---a stark contradiction to the celebrated \unix philosophy.
For example, \ttt{parallel} contains flags such as \ttt{--skip-first-line}, \ttt{-trim}, and \ttt{--xargs}, that a \unix user can achieve using \ttt{head}, \ttt{sed}, and \ttt{xargs}; it also includes other programs with complex semantics, such as the ability transfer files between computers, separate text files, and parse CSV.
\sys embraces the \unix philosophy and simply attempts to rewrite pipelines more efficiently.

Several shells~\cite{duff1990rc, mcdonald1988support, dagsh:17} add primitives for non-linear pipe topologies---some of which target distribution.
Here too, however, developers are expected to manually rewrite scripts to exploit these new primitives.

@@

\heading{Cluster Computing}
Modern analytics frameworks~\cite{mapreduce:08, ciel:11, spark:12, naiad:13} and domain-specific languages~\cite{alvaro2011consistency, distal:13, meiklejohn2015lasp} go a step further in terms of automation by providing a few carefully-designed primitives.
These primitives make strong assumptions about the nature of the computation---\eg strongly-eventual commutative functions that can proceed in parallel.
By targeting specific classes of computation (\emph{viz.} \sys's distributability), they render distributed computation efficiently.

% Lack of generality, no formal guaratees, their setup and is too tedious 
\sys chooses a more general approach:
  it does not require users to set up a new framework for every new class of computation they need to execute, nor to rewrite different parts of their computation using the specific abstractions provided by each framework they use.
It also takes a more principled approach: by developing the semantic model outlined earlier, it can guarantee that transformations are correct with respect to the sequential program.

% % Although still a research prototype, \sys's setup and deployment is simpler.
% It also does not require these specialized frameworks, manually (re)write , and 
% Different from \sys, however, 
% Due to the specialization of these primitives, however, these systems suffer from a lack of generality ;
% % and they still require developers to compose their work using the pipelines
% Developing under these frameworks differs quite significantly from the development of normal (non-distributed) programs.

\heading{Annotation-based Distribution}
Recent systems on light-touch distribution, such as Ignis~\cite{ignis:19} and Mozart~\cite{mozart:19}, are more general than cluster-computing frameworks.
They also lower (but not entirely eliminate) developer effort by automating most of the rewriting process.
Both work on high-level, managed languages---server-side JavaScript for Ignis and Python for Mozart---that allow runtime transformations without affecting the broader environment under which a program executes.
For the critical points where user input is required, they provide a high-level declarative DSL with varying (but limited) degrees of expressive power.
% Annotations are used by ---complicated by the use of third-party libraries similar to \sys's extensions.
Unfortunately, a challenge with this approach is that users composing applications by using third-party libraries do not understand their distributability details well-enough to annotate them.

\sys occupies a different point in the design space, by removing any work from composers and adding a small load to developers.
% More specifically, it is designed to be used automatically by composers,
Its DSL is designed to be more heavy-weight than Ignis' declarative recipes or Mozart's split annotations, but geared towards the developers of these components---not the users composing them.

% It also provides a more principled approach than both of these systems.
% Their annotations language is not as principled as \sys's---and one that is critical in ensuring that the dataflow analyses preserves the semantics expected by the developer
% The side-effects present in the shell dwarf those of functional languages

% Nice intro: http://homepages.inf.ed.ac.uk/bfranke/Publications/pldi121-tournavitis.pdf
\heading{Low-level Parallelization}
Instruction-level parallelization has a longer history, starting from explicit \ttt{DOALL} and \ttt{DOACROSS} annotations~\cite{par1, par2} and continuing with compilers that attempt to automatically extract parallelism~\cite{padua1993polaris,hall1996maximizing}.
These systems operate at a lower level than \sys (\eg that of instructions or loops instead of pipe boundaries) and typically do not exploit runtime information.

% https://ecommons.cornell.edu/bitstream/handle/1813/6508/85-668.pdf?sequence=1
% The growth of the web led to specialized frameworks for massively distributed computation~\cite{mapreduce:08, spark:10, naiad:13}.
% While they Moreover, these systems take advantage of functional purity;
%   for programs that are not data-intensive processing pipelines (\eg web servers), purely functional code is generally responsible for only a small fraction of the program runtime.
